\documentclass[11pt]{article}
\usepackage{cv_style}
\usepackage{multicol}

\begin{document}

\author{Shenxing Zhang}
\address{B1810 Jade Science and Education Building, 485 Danxia Road, Hefei Anhui PRC\par Hefei University of Technology}
\email{zhangshenxing@hfut.edu.cn}
\website{https://zhangshenxing.github.io/}
\maketitle

\section{Education}
\begin{cvstage}
	\editem{Ph.D. Mathematics}
		{Nov. 2015}
		{School of Mathematical Sciences, University of Science and Technology of China}
		{Advisor: Yi Ouyang}
	\editem{B.A. Mathematics}
		{Jun. 2010}
		{School of the Gifted Young, University of Science and Technology of China}
		{Advisor: Yi Ouyang}
	\editem{Visiting Student}
		{May 2015 -- Jun. 2015}
		{Department of Mathematics, University of California,  Berkeley, USA}
		{Advisor: Xinyi Yuan}
\end{cvstage}


\section{Academic Appointment}
\begin{cvstage}
	\editem{Research Associate Professor}
		{Dec. 2021 -- Present}
		{School of Mathematics, Hefei University of Technology}
		{}
	\editem{Postdoctoral Fellow}
		{Apr. 2018 -- Nov. 2021}
		{School of Mathematical Sciences, University of Science and Technology of China}
		{Mentor: Yi Ouyang}
	\editem{Postdoctoral Fellow}
		{Mar. 2016 -- Feb. 2018}
		{Academy of Mathematics and Systems Science, Chinese Academy of Sciences}
		{Mentor: Ye Tian}
\end{cvstage}


\section{Publications and Preprints}
\subsection{Publications}
\begin{cvlist}
	\item \textbf{S. Zhang}.
	The generating fields of twisted Kloosterman sums.
	{\em Int. J. Number Theory}, (2025), to appear.

	\item \textbf{S. Zhang}.
	On the linearity of the periods of subtraction games.
	{\em Theor. Comput. Sci.}, 985 (2024).
	
	\item \textbf{S. Zhang}.
	On a comparison of Cassels pairings of different elliptic curves.
	{\em Acta Arith.} 211 (2023), no. 1, 1--23.
	
	\item \textbf{S. Zhang}.
	On the Newton polygons of twisted $L$-functions of binomials.
	{\em Finite Fields Appl.} 80 (2022), Paper No. 102026, 20 pp.

	\item J. Li, \textbf{S. Zhang}.
	The $3$-class groups of $\mathbb{Q}(\sqrt[3]{p})$ and its normal closure.
	{\em Math. Z.} 300 (2022), no. 1, 209--215.

	\item J. Li, Y. Ouyang, Y. Xu, \textbf{S. Zhang}.
	$\ell$-Class groups of fields in Kummer towers.
	{\em Publ. Mat.} 66(2022), no. 1, 235--267.

	\item \textbf{S. Zhang}.
	The generating fields of two twisted Kloosterman sums.
	{\em J. Univ. Sci. Technol. China} 51 (2021), no. 12, 879-888.

	\item Y. Ouyang, \textbf{S. Zhang}.
	Birch's lemma over global function fields.
	{\em Proc. Amer. Math. Soc.} 145 (2017), no. 2, 577--584.

	\item Y. Ouyang, \textbf{S. Zhang}.
	Newton polygons of $L$-functions of polynomials $x^d+ax^{d-1}$ with $p\equiv -1 \bmod d$.
	{\em Finite Fields Appl.} 37 (2016), 285--294.

	\item Y. Ouyang, \textbf{S. Zhang}.
	On second 2-descent and non-congruent numbers.
	{\em Acta Arith.} 170 (2015), no. 4, 343--360.

	\item Y. Ouyang, \textbf{S. Zhang}.
	On non-congruent numbers with 1 modulo 4 prime factors.
	{\em Sci. China Math.} 57 (2014), no. 3, 649--658.
\end{cvlist}


\subsection{Preprints}
\begin{cvlist}
	% \item \textbf{S. Zhang}.
	% On the Diophantine equation $my^2=x^3-k^3$.
	% (2025), submitted.

	\item Z. Wang, \textbf{S. Zhang}.
	On the quadratic twist of elliptic curves with full $2$-torsion.
	(2022), submitted.

	\item \textbf{S. Zhang}.
	The virtual period of the degree sequences of the exponential sums.
	{\em Arxiv:} 2010.08342, preprint.
\end{cvlist}


\section{Grants \& Awards}
\subsection{Grants}
\begin{cvlist}
	\item 173 Program, Co-PI
	\item Fundamental Research Funds for the Central Universities (JZ2023HGTB0217)
	\item NSFC (12001510)
	\item China Postdoctoral Science Foundation (2017M611027)
	\item Fundamental Research Funds for the Central Universities (WK0010000061), Co-PI
	\item NSFC (12271335, 11601255, 11571328, 11201445, 11171317), Co-PI
\end{cvlist}


\subsection{Awards}
\begin{cvlist}
	\item Young Teachers Teaching Basic Skills Competition in HFUT, 2nd Prize, 2023.
	\item National Scholarship for Graduate Students, 2013
	\item National Scholarship for Undergraduate Students, 2009
	\item Anhui Excellent Young Student, 2006
\end{cvlist}


\section{Teaching Experience}
\subsection{Teaching}
\begin{cvstage}
	\item Hefei University of Technology
	\begin{cvstage}
		\csitem{1400261B Complex Function and Integral Transformation}{Fall 2022-2024}
		\csitem{1400071B Linear Algebra}{Fall 2024}
		\csitem{034Y01 Mathematics (II)}{Spring 2022-2023}
	\end{cvstage}
	\item University of Science and Technology of China
	\begin{cvstage}
		\csitem{001548 Complex Variables B}{Fall 2020}
		\csitem{MA05109 Algebraic Number Theory}{Spring 2020}
	\end{cvstage}
\end{cvstage}


\subsection{Teaching Assistant}
\begin{cvstage}
	\item Tsinghua University
	\begin{cvstage}
		\csitem{The Yau Tsinghua Mathcamp, Linear Algebra}{Summer 2021-2023}
	\end{cvstage}
	\item Hefei University of Technology
	\begin{cvstage}
		\csitem{Probability and Statistics}{Fall 2021}
	\end{cvstage}
	\item University of Chinese Academy of Sciences
	\begin{cvstage}
		\csitem{Summer School on Algebra and Number Theory}{Summer 2019}
		\csitem{Basic Number Theory}{Fall 2016}
	\end{cvstage}
	\item University of Science and Technology of China
	\begin{cvstage}
		\csitem{001356 Basic Algebra}{Fall 2014}
		\csitem{001704 Abstract Algebra (Honor)}{Spring 2013}
		\csitem{001356 Basic Algebra}{Fall 2012}
		\csitem{001010 Abstract Algebra}{Fall 2011}
		\csitem{001012 Complex Variables}{Spring 2011}
	\end{cvstage}
\end{cvstage}


\subsection{Publications on Teaching}
\begin{cvlist}
	\item Z. Yuan, \textbf{S. Zhang}.
	Research on Laurent series of complex functions in different annulus (in Chinese).
	{\em Journal of Nanyang Normal University}, 2024, 23(3): 38--41.
	\item Y. Ouyang, \textbf{S. Zhang}.
	Theoretic application of elementary transformations in linear algebra teaching (in Chinese).
	{\em College Mathematics}, 2023, 39(5): 68--75.
\end{cvlist}


\subsection{Advisers}
\begin{cvstage}
	\item Bachelor's Thesis
	\begin{cvstage}
		\csitem{X. Qiu, Research on the isogenies of elliptic curves and its applications on cryptography}{2024}
		\csitem{X. Liu, The class groups of quadratic fields and their applications}{2022}
	\end{cvstage}
	\item National Student Innovation and Entrepreneurship Training Program
	\begin{cvstage}
		\csitem{S. Zhu, Y. Wang, Y. Deng, J. Yao, T. Luo, A contagion model of 2019-nCoV}{2022-2023}
	\end{cvstage}
\end{cvstage}


\section{Referee for Journals}
\begin{multicols}{2}
	\begin{cvlist}
		\item {\em International Journal of Number theory}
		\item {\em Algebra Colloquium}
		\item {\em The Ramanujan Journal}
		\item {\em Finite Fields and Their Applications}
		\item {\em Pure and Applied Mathematics Quarterly}
		\item {\em Bulletin of the Belgian Mathematical Society – Simon Stevin}
		\item {\em Scientia Sinica Mathematica (Chinese)}
		\item {\em College Mathematics (Chinese)}
	\end{cvlist}
\end{multicols}


\section{Research Interests}
\begin{multicols}{2}
	\begin{cvlist}
		\item Elliptic curves
		\item Exponential sums
		\item Class groups
		\item Iwasawa theory
		% \item $p$-adic Hodge theory
	\end{cvlist}
\end{multicols}

\medskip\emph{Last updated in March 29\textsuperscript{th}, 2025}

\end{document}
